---
title: "**【1204】**"
output: 
  html_document:
    df_print: paged
---

**$\mathcal{YiHsin}\;\mathcal{Lu}$**

****

# **1**

## **The attached data `iqsize.txt` including four variables performance IQ scores PIQ, brain size from MRI scans Size (count/10,000), height Height (inches), and weight weight (pounds) for n=38 college students.**

```{r echo=FALSE}
load('1204.Rdata')
```

****

## (a) Please separately calculate the 95 % confidence region of new respondent’s IQ score with the given data.

### **The given data:**
```{r echo=FALSE}
as.data.frame(x_h[,-1])
```

### **The ($1-\alpha$)% Confidence Interval:**

$$
\Large{\left (\;\;\hat{Y}_h\pm t_{(1-\alpha/2, n-p)}\sqrt{\mbox{MSE}(1+\mathbb{X}_h(\mathbb{X}'\mathbb{X})^{-1}\mathbb{X'}_h)}\;\;\right )}
$$

### **The 95% Confidence Interval:**

$$
\Large{\left (\;\;\hat{Y}_h\pm t_{(0.975, 34)}\sqrt{\mbox{MSE}(1+\mathbb{X}_h(\mathbb{X}'\mathbb{X})^{-1}\mathbb{X'}_h)}\;\;\right )}
$$

```{r echo=FALSE}
as.data.frame(ci1)
```

****

## (b) Consider simultaneous
confidence region by Bonferroni’s method.

### **The 95% Confidence Interval:**

$$
\Large{\left (\;\;\hat{Y}_h\pm t_{(1-\frac{0.05}{2\cdot 7}, 34)}\sqrt{\mbox{MSE}\cdot \mathbb{X}_h(\mathbb{X}'\mathbb{X})^{-1}\mathbb{X'}_h}\;\;\right )}
$$

```{r echo=FALSE}
as.data.frame(ci2)
```

****

## (c) Compare the length of constructed confidence region in (a) and (b)

```{r echo=FALSE}
lci
```