---
title: "**【APLM-Midterm 5】**"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

**$\mathcal{YiHsin}\;\mathcal{Lu}$**

****

```{r include=FALSE}
library(ggplot2)
library(GGally)
load('1110.Rdata')
```

# **5. **

Based on the given data, the variables are defined as Performance IQ scores (PIQ)
from the revised Wechsler Adult Intelligence Scale. Brain: Brain size based on
the count obtained from MRI scans (given as count/10,000). Height: Height in
inches. Weight: Weight in pounds. (Analysis report and the related codes are
required. Please present the programming code as the attachment.)


****

## **【Linear Model】**

### $y_i=\beta_0+\beta_1x_{i1}+\beta_2x_{i2}+\beta_3x_{i3}+\epsilon_i\;,\;\forall\; i=1,\cdots,38$

* $\large{y_i}$: PIQ
* $\large{x_{i1}}$: Brain
* $\large{x_{i2}}$: Height
* $\large{x_{i3}}$: Weight


```{r echo=FALSE}
iq
```

```{r}
summary(lm_iq)
```

****

## **【Pairs Plot】**

```{r echo=FALSE}
ggpairs(iq)
```

```{r}
newdata1 = as.data.frame(matrix(data = c(70,60,120,
                                        90,90,50), ncol = 3, nrow = 2, byrow = T))
colnames(newdata1) = c('Brain','Height','Weight')
predict(lm_iq, newdata = newdata1)
```