---
title: "**【APLM:assignment1031】**"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

**$\mathcal{YiHsin}\;\mathcal{Lu}$**

****

```{r include=FALSE}
library(ggplot2)
library(GGally)
load('1110.Rdata')
```

# **C. **

The attached data `iqsize.txt` including four variables performance IQ scores PIQ, brain size from MRI scans Size (count/10,000), height Height (inches), and weight (pounds) for n=38 college students. Analyze the data and make a summary.In addition, check whether the mentioned physical information affect IQ score performance.

****

## **【Linear Model】**

### $y_i=\beta_0+\beta_1x_{i1}+\beta_2x_{i2}+\beta_3x_{i3}+\epsilon_i\;,\;\forall\; i=1,\cdots,38$

* $\large{y_i}$: PIQ
* $\large{x_{i1}}$: Brain
* $\large{x_{i2}}$: Height
* $\large{x_{i3}}$: Weight


```{r echo=FALSE}
iq
```

```{r}
summary(lm_iq)
```

****

## **【Pairs Plot】**

```{r echo=FALSE}
ggpairs(iq)
```

## **【T-test】**

### **1. $\beta_1 = \cfrac{Y-\hat{\beta_0}-\hat{\beta_2}X_2-\hat{\beta_3}X_3}{X_1}$**

* $\Large{H_0:\bar{\beta_1}=0}$


* $\Large{H_1:\bar{\beta_1}\neq0}$

```{r echo=FALSE}
t.test(beta1/x1, mu=0)
```

****

### **2. $\beta_2 = \cfrac{Y-\hat{\beta_0}-\hat{\beta_1}X_1-\hat{\beta_3}X_3}{X_2}$**

* $\Large{H_0:\bar{\beta_2}=0}$


* $\Large{H_1:\bar{\beta_2}\neq0}$

```{r echo=FALSE}
t.test(beta2/x2, mu=0)
```

****

### **3. $\beta_3 = \cfrac{Y-\hat{\beta_0}-\hat{\beta_1}X_1-\hat{\beta_2}X_2}{X_3}$**

* $\Large{H_0:\bar{\beta_3}=0}$


* $\Large{H_1:\bar{\beta_3}\neq0}$

```{r echo=FALSE}
t.test(beta3/x3, mu=0)
```