---
title: "**【APLM: homework0929】**"
output: 
  html_document:
    df_print: paged
---

**$\mathcal{YiHsin}\;\mathcal{Lu}$**

****

  
```{r include=FALSE}
library(plotly)
library(lattice)
library(matlib)
load("D:/1006.RData")
```


## **【EX1】**
We would like to realize which factors are related to the likelihood that a hospital patients acquires an infection while hospitalized. The variables here are y= infection risk, x1= average length of patient stay, x2= average patient age, x3= measure of how many x-rays are given in the hospital (Hospital Infection dataset). 

### **(EX1) Data Set**

```{r echo=FALSE}
hosData %>% select(4, 2, 3, 6)
```

****

### **(EX1) Least Square Estimation**

$\hat\beta = (\mathbb{X}'\mathbb{X})^{-1}\mathbb{X}'\mathbb{Y}$

```{r echo=FALSE}
rownames(lse_hosData$beta_hat) = c('beta0_hat', 'beta1_hat', 'beta2_hat', 'beta3_hat')
colnames(lse_hosData$beta_hat) = 'estimator'
lse_hosData$beta_hat
```

****

## **【EX2】**
There are 214 females in statistics classes at the University of California at Davis (Stat Females dataset). The variables are y= student’s self-reported height, x1= student’s guess at her mother’s height, and x2= student’s guess at her father’s height. All heights are in inches.

### **(EX2) Data Set**

```{r echo=FALSE}
heiData
```

****

### **(EX2) Least Square Estimation**

* $\hat\beta = (\mathbb{X}'\mathbb{X})^{-1}\mathbb{X}'\mathbb{Y}$

```{r echo=FALSE}
rownames(lse_heiData$beta_hat) = c('beta0_hat', 'beta1_hat', 'beta2_hat')
colnames(lse_heiData$beta_hat) = 'estimator'
lse_heiData$beta_hat
```

****

### **(EX2) Plot**

* Surface: $\hat{y} = \hat\beta_{0}+\hat\beta_{1}x_1+\hat\beta_{2}x_2$

* Red points: $(x_1,x_2,y)$

```{r echo=FALSE}
fig
```

## **【assigment0930(d.)】**

### **1 $E(\epsilon)=0$**

* $Y_i = \beta_{0}+\beta_{1}X_i+\epsilon_i$

* $\beta_{0} = 0.3$

* $\beta_{1} = 2$

* $\epsilon_i\sim\mathbb{N}(0, \sigma^2)$

#### **1.1 estimation for $\;\beta$**

```{r echo=FALSE}
beta
```

#### **1.2 Plot**

```{r echo=FALSE}
reg0var
```

### **2 $E(\epsilon)\neq 0$**

* $Y_i = \beta_{0}+\beta_{1}X_i+\epsilon_i$

* $\beta_{0} = 0.3$

* $\beta_{1} = 2$

* $\epsilon_i\sim\mathbb{N}(2, \sigma^2)$

#### **2.1 estimation for $\;\beta$**

```{r echo=FALSE}
beta1
```

#### **2.2 Plot**

```{r echo=FALSE}
reg2var
```